---
title: "Guided Notes 1.1: Introduction to Data Visualization"
author: "Your Name "
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, fig.height = 4, fig.width = 6)
```

These notes correspond to material from Chapter 3 of R for [Data Science](https://r4ds.had.co.nz) by Wickham and Grolemund.


## Loading and Exploring Data

#### Loading Data and R packages

We'll load an R package called `tidyverse`. This is a poweful R package that is useful for creating graphics, and wrangling data, and many other tasks.  

If the package is not already installed, you will need to install it, using the instructions in Assignment 0. Once installed, we load the package using the `library()` command. You will need to do this everytime you start a new R session and want to use the package. 

```{r}
library(tidyverse)
```

We will work with a dataset containing information on movies released in 2018. Data were collected from https://www.imdb.com/. 

We first read in the data from my github repository, where it is posted, and assign it the name "movies"

```{r}
movies <- read_csv("https://raw.githubusercontent.com/AndrewjSage/andrewjsage.github.io/master/Datasets/movies2018.csv")
```

#### Exploring the Data

The `head()` command displays the first 6 rows of the dataset

```{r}
head(movies)
```

The `glimpse()` command returns a summary of the information contained in the dataset, including variable types. 

```{r}
glimpse(movies)
```

* Categorical variables take on groups or categories, rather than numeric values. In R, these might be coded as logical `<logi>`, character `<chr>`, factor `<fct>` and ordered factor `<ord>`.  


* Quantitative variables take on meaningful numeric values. These include numeric `<num>`, integer `<int>`, and double `<dbl>`.   

The `summary()` command returns a summary of each variable in the dataset

```{r}
summary(movies)
```

##  Creating Graphics 

The `ggplot` command is used to create graphics in `tidyverse`. 

### Example 1.1.1 Basic Scatterplot

We create a scatterplot of IMDB score vs. runtime for 2018 movies. 

```{r}
ggplot(data = movies) + 
  geom_point(mapping = aes(x = run_time, y = IMDB_score))
```

### Example 1.1.2 Scatterplot with text labels

Out of curiousity, let's add the names of the films that scores especially high or low. 

We'll see the `filter` command more later on. 

```{r, fig.height=6}
ggplot(data = movies) + 
  geom_point(mapping = aes(x = run_time, y = IMDB_score)) +
  geom_text(data=movies %>% filter(IMDB_score > 8 | IMDB_score < 4 ), 
            aes(x = run_time, y = IMDB_score, label = title, vjust=-1), 
            size = 2, color="red")
```

### Example 1.1.3 Scatterplot with color, shape, and size


We can use color, shape, and size to convey information about other variables. 

We'll use color for genre, shape for MPAA rating, and size for gross revenue generated.

Note that shape should be used for factor variables, and size for numeric variables. Color may be used for either. 

```{r}
ggplot(data = movies) + 
  geom_point(mapping = aes(x = run_time, y = IMDB_score, 
                           color=genre, shape=MPAA_rating, size=gross))
```


### Example 1.1.4 Changing Axis Labels and Hiding Legend

```{r}
ggplot(data = movies) + 
  geom_point(mapping = aes(x = run_time, y = IMDB_score, 
                           color=genre, shape=MPAA_rating, size=gross)) +
  xlab("Runtime (in minutes)")+ ylab("IMDB Score (out of 10)")
```

#### Practice 1.1.5 

Fill in the blanks to create a scatterplot that shows IMDB score on the y-axis and critics score (`meta_score`) on the x-axis. Use color to denote MPAA_rating, and size to represent the number of votes on IMDB (`vote`). Add appropriate axis lables. 

When you are done, remove `eval=FALSE` from the braces in the chunk heading, so your code will run. 

```{r, eval=FALSE}
ggplot(data = ___________) + 
  geom_point(mapping = aes(x = ____________, y = ___________, 
                           color=__________, size=________)) +
  xlab("_____________")+ylab("__________")
```


