---
title: "Guided Notes 1.3: Advanced Data Visualization"
author: "Your Name"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, fig.height = 4, fig.width = 6)
```


## Advanced Data Graphics

Let's take a look at some graphics that go beyond the basics. We'll recreate two graphics that orignally appeared in a fivethirtyeight.com blogpost [How to tell Someone's Age when All You Know is Her Name](https://fivethirtyeight.com/features/how-to-tell-someones-age-when-all-you-know-is-her-name/)  

This example is heavily based on one from Section 3.3 of [Modern Data Science with R](https://beanumber.github.io/mdsr2e/), by Baumer, Kaplan, and Horton. It has been modified slightly. 

```{r}
library(tidyverse)
library(babynames)
library(mdsr)
```

### Babynames dataset 

```{r}
Babynames_data <- make_babynames_dist()
head(Babynames_data)
```

The data are broken down by gender (M, F). The following step produces a dataset with all babies of a given name, regardless of gender. 

We'll talk more about the commands being used soon, but for now, you can just run the code and use the resulting dataset. 

```{r}
Babynames_data2 <- Babynames_data %>% group_by(year, name) %>% summarize(n=sum(n), alive_prob=mean(alive_prob), age_today=mean(age_today)) %>% mutate(est_alive_today = round(n*alive_prob,0), count_thousands = n/1000)
head(Babynames_data2)
```

### Example 1.3.1: Age Distribution for a Given Name  

### Choosing Name 

```{r}
Babyname <- Babynames_data2 %>% filter(name == "Joseph")

name_plot <- ggplot(data=Babyname, aes(x=year)) + 
  geom_bar(aes(y=est_alive_today/1000), fill="#b2d7e9", color="white", stat="identity")
name_plot
```

See the [R Cookbook](http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/) for more information on color palette. 

### Setting Axis Limits 

```{r}
name_plot <- name_plot + ylim(c(0,42))
name_plot
```

### Adding Line  

```{r}
name_plot <- name_plot + geom_line(aes(y=count_thousands), size=2)
name_plot
```

### Axis Labels 

```{r}
name_plot <- name_plot + 
  xlab("Year") + ylab("Number of people (in thousands)")
name_plot
```

### Adding Text 

```{r}
name_plot <- name_plot + 
  ggtitle("Age Distribution of Americans Named Joseph") + 
  geom_text(x=1935, y=40, label="Number of Josephs\nborn each year") +
  geom_text(x=1915, y=13, label="Number of Josephs \nborn each year\nestimated to be alive on \n 1/1/2014", color="#b2d7e9") 
name_plot
```


### Adding Median Age  

```{r}
#install.packages("Hmisc")

library(Hmisc)
#wtd.quantile <- Hmisc::wtd.quantile
median_yob <- with(Babyname, wtd.quantile(year, est_alive_today, probs=0.5))
median_yob
```

```{r}
name_plot <- name_plot + 
  geom_bar(stat="identity", color="white", fill="#008fd5", aes(y = ifelse(year==median_yob, est_alive_today / 1000 , 0)))
name_plot
```

```{r}
name_plot <- name_plot + 
  geom_text(x=2003, y=40, label="The median \nliving Joseph \nis 37 years old", color="darkgray") + 
  geom_curve(x=1995, xend=1974, y=40, yend=24, arrow=arrow(length=unit(0.3, "cm")), curvature=0.5) 
name_plot
```

### Example 1.3.2: Common Female Names  

### Some data prep  

```{r}
common_names <- Babynames_data %>%
  # hack added in 2nd edition
  filter(n > 100) %>%
  filter(sex == "F") %>% 
  group_by(name) %>%
  mutate(wgt = est_alive_today / sum(est_alive_today)) %>%
  summarise(
    N = n(), est_num_alive = sum(est_alive_today),
    q1_age = wtd.quantile(age_today, est_alive_today, probs = 0.25, na.rm = TRUE),
    median_age = wtd.quantile(age_today, est_alive_today, probs = 0.5, na.rm = TRUE),
    q3_age = wtd.quantile(age_today, est_alive_today, probs = 0.75, na.rm = TRUE)
  ) %>%
  arrange(desc(est_num_alive)) %>%
  head(25)
```

### Setup the Plot  

We put the names along the x-axis in decreasing order of median age.

```{r}
w_plot <- ggplot(data = common_names, 
                 aes(x = reorder(name, -median_age), 
  y = median_age)) + 
  xlab(NULL) + ylab("Age (in years)") + 
  ggtitle("Median ages for females with the 25 most common names")
w_plot
```

### Add Linerange and color  

```{r}
w_plot <- w_plot + 
  geom_linerange(aes(ymin = q1_age, ymax = q3_age), 
  color = "#f3d478", size = 10, alpha = 0.8)
w_plot
```

### Add points at median  

```{r}
w_plot <- w_plot +
  geom_point(fill = "#ed3324", colour = "white", size = 4, shape = 21)
w_plot
```

### Add Labels and Flip Coordinates  

```{r, fig.height=7, fig.width=7}  
w_plot <- w_plot + 
  geom_point(aes(y = 55, x = 24), fill = "#ed3324", colour = "white", 
    size = 4, shape = 21) + 
  geom_text(aes(y = 58, x = 24, label = "median")) + 
  geom_text(aes(y = 26, x = 16, label = "25th")) + 
  geom_text(aes(y = 51, x = 16, label = "75th percentile")) + 
  geom_point(aes(y = 24, x = 16), shape = 17) + 
  geom_point(aes(y = 56, x = 16), shape = 17) +
  coord_flip()  
w_plot
```